---
title: Loading multiple *.csv (*.xlsx) effectively
author: Petr Schönbauer
date: '2017-03-08'
slug: loading-multiple-csv-xlsx-effectively
categories: []
tags:
  - ggplot2
  - map
  - R
description: Desc
hacker_news_id: ''
lobsters_id: ''
meta_img: /images/image.jpg
---

This brief post is a "how to" load multiple `csv` / `xlsx` files into single data frame. Say, we have a couple of *.xlsx files in one folder (some of them in sub-directories) and we would like to bind them all into one. Quite common task. Let´s do it:)

```{r, message=FALSE}
library(tidyverse)
library(readxl)
```

# Step 1
`list.files` function with argument `full.names = TRUE` (preserving complete path to all files) and argument `recursive = TRUE` is handy in case your files are in any level of sub-directories (represented with "next_level" folder in my case).

```{r}
x <- list.files(path = "../../static/data/multiple", full.names = TRUE, recursive = TRUE, pattern = "*.xlsx")
x
```

Now I have the `x` vector with all full paths to all "*.xlsx" files.

# Step 2
`map` the vector of paths to function `read_excel`. Use argument `sheets` to specify the name of the sheet.

```{r}
lst <- map(x, read_excel, sheet = "Value Assessment")
head(lst[[1]])
```

These are first 5 records of first file.

# Step 3
Use `bind_rows` to bind the list into one tibble (data frame). Argument `.id` allows you to create a column with the name of source file.

```{r}
df <- bind_rows(lst, .id = "filename")
head(df)
```

# Pipe pipe pipe

Let´s pipe all these steps into single chunk of code.

```{r}
list.files(path = "../../static/data/multiple", full.names = TRUE, recursive = TRUE, pattern = "*.xlsx") %>% 
  map(read_excel, sheet = "Value Assessment") %>% 
  bind_rows(.id = "filename") %>% 
  head
```

Actually, you can reduce it to two lines using `map_df`:)

```{r}
list.files(path = "../../static/data/multiple", full.names = TRUE, recursive = TRUE, pattern = "*.xlsx") %>% 
  map_df(~read_excel(., sheet = "Value Assessment"), .id = "filename") %>% 
  head
```

Pipes are cool, makes your life easier. Two lines of code, but very powerfull;)
