---
title: Digit recognition with TensorFlow
author: Petr Schönbauer
date: '2018-11-19'
slug: digit-recognition-with-tensorflow
categories: []
tags:
  - R
  - tensorflow
  - machine learning
draft: true
---

Introduction myself to machine learing with Keras / TensorFlow. This post is based on [R Studio Keras blog tutorial](https://tensorflow.rstudio.com/keras/)

# What is what?

* Keras = high-level neural networks API for TensorFlow written in Python
* TensorFlow = Python-friendly open source library for numerical computation
* Anaconda = Python distribution (platform for data science)

Ok, Python, Python, Python - I think I´ll need Python:) to be installed besides R package Keras. It will not be that easy. 

First step - install ["Anaconda ditribution"](https://www.anaconda.com/download/) - should play well with Keras (Stackoverflow masters say so, so it must be truth:)).

# R
```{r, message=FALSE}
library(keras)
library(tidyverse)
library(tensorflow)
```

Install Keras first. By default the CPU-based installation of Keras and TensorFlow is installed. Seems I can use GPU based one too, maybe later. 
```{r, message=FALSE}
# install_keras()
```

# MNIST digit recognition aka machine learning "Hello world"
MNIST = Modified National Institute of Standards and Technology database is a large database of handwritten digits that is commonly used for training various image processing systems. `dataset_mnist()` function will download the dataset.
```{r}
mnist <- dataset_mnist()
```

```{r}
mnist %>% glimpse
```

Dataset contains two lists - train and test set. 

* `train` contains 60000 observations (images) for training
* `test` contains 10000 observation (images) for testing.

* `x` is a 3D array - no. of image, width, height
* `y` is a 1D array - corresponding digit (class)

3D array is basically set of matrices. Each image is described by 28x28 matrix. Each point represent pixel. The value represent color in greyscale range of 0-255.

For example, code below indexes first 20x10 pixels of 3rd image from `train` set. We can recognize piece of digit 4. 

```{r}
mnist$train$x[3, 1:20, 1:10]
mnist$train$y[3]
```

```{r}
x_train <- mnist$train$x
y_train <- mnist$train$y
x_test <- mnist$test$x
y_test <- mnist$test$y
```

```{r}
# reshape
x_train <- array_reshape(x_train, c(nrow(x_train), 784))
x_test <- array_reshape(x_test, c(nrow(x_test), 784))
# rescale
x_train <- x_train / 255
x_test <- x_test / 255
```

```{r}
y_train <- to_categorical(y_train, 10)
y_test <- to_categorical(y_test, 10)
```

```{r}
model <- keras_model_sequential() 
model %>% 
  layer_dense(units = 256, activation = 'relu', input_shape = c(784)) %>% 
  layer_dropout(rate = 0.4) %>% 
  layer_dense(units = 128, activation = 'relu') %>%
  layer_dropout(rate = 0.3) %>%
  layer_dense(units = 10, activation = 'softmax')
```

```{r}
model %>% compile(
  loss = 'categorical_crossentropy',
  optimizer = optimizer_rmsprop(),
  metrics = c('accuracy')
)
```

```{r}
history <- model %>% fit(
  x_train, y_train, 
  epochs = 30, batch_size = 128, 
  validation_split = 0.2
)
```

```{r}
plot(history)
```


```{r}
model %>% evaluate(x_test, y_test)
```

```{r}
model %>% predict_classes(x_test) %>% head
```

