---
title: "Top 20 movies based on IMDB scores"
author: "Petr Schönbauer"
date: "2017-09-10"
slug: top-20-movies-based-on-imdb-scores
tags:
- R
- IMBD
- ggplot2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=TRUE, message=FALSE, warning=FALSE}
# load libraries
library(tidyverse)
library(shiny)
```

# Basic data wrangling first

## Load and check the data structure 

```{r, message=FALSE, warning=FALSE}
# load data
raw <- read_csv("../../static/data/movie_data.csv")
# structure of data
glimpse(raw)
# show character and numeric variables separately
raw %>% 
  select_if(is.character) %>% 
  glimpse
raw %>% 
  select_if(is.numeric) %>% 
  glimpse
```

## Trim the `movie_title` variable

Types of variables seems to be reasonably assigned. For some reason there ia a whitespace at the end of each `movie_title`. Let´s fix it with `str_trim` from `stringr` library.

```{r}
# remove whitespaces from movie_title
df <- raw %>% 
  mutate(movie_title = str_trim(movie_title))

df$movie_title %>% head
```

No more whitespaces.

## Remove duplicated movies

Are there some duplicated movies?

```{r}
# count of observations (movies)
nrow(df)

# count of unique movies
df %>% 
  distinct(movie_title) %>% 
  count()

# or using dplyr´s n_distinct function
n_distinct(df$movie_title)
```

Alternative:

```{r}
# any duplicated?
any(duplicated(df$movie_title))

# sum of duplicated movies
sum(duplicated(df$movie_title))
```
Yes, there are some, let´s exclude them.

```{r}
# remove duplicates
df <- df[!duplicated(df$movie_title),]

# count of unique movies after removing of duplicates
df %>% 
  distinct(movie_title) %>% 
  count()
```

# Simple exploratory analysis

## Top 20 movies according to IMDB scores

First check the `num_voted_users` of movies with top 10 `imdb_scores`.

```{r}
df %>% 
  arrange(desc(imdb_score)) %>% 
  slice(1:20) %>% 
  arrange(num_voted_users) %>%
  select(movie_title,num_voted_users) %>% 
  print(n=10)
```

Obviously, there is some issue in data. For example only 10 voters for movie "Towering inferno" seems not to be plausible. Assuming the IMDB scores are correctly collected let´s exclude the movies with less than 10 000 votes. It seems to be resonable threshold. In addition color the bars by country of origin.

```{r}
# plot the top 20 movies and color by country
df %>%
  filter(num_voted_users > 10000) %>% 
  arrange(desc(imdb_score)) %>% 
  slice(1:20) %>% 
  ggplot(aes(reorder(movie_title, imdb_score), imdb_score, fill = country)) +
  geom_bar(stat = "identity") +
  scale_y_continuous(breaks = seq(0,10,1)) +
  coord_flip() +
  labs(title = "Top 20 movies according to IMDB scores",
       subtitle = "by country",
       y = "IMDB score",
       x = "Movie title") +
  theme(plot.title = element_text(face = "bold",
                                  size = 16,
                                  hjust = 1),
        plot.subtitle = element_text(face = "italic",
                                     size = 11,
                                     hjust = 0)
  )
                                     
```

As expected, most of the movies were produced in US.

## Top n movies by country

Say, I´m interested in top 10 movies from UK only.

```{r}
# plot the top 10 movies from UK
df %>% 
  filter(num_voted_users > 10000) %>% 
  filter(country %in% c("UK")) %>% 
  arrange(desc(imdb_score)) %>% 
  slice(1:10) %>% 
  ggplot(aes(reorder(movie_title, imdb_score), imdb_score, fill = country)) +
  geom_bar(stat = "identity") +
  scale_y_continuous(breaks = seq(0,10,1)) +
  coord_flip() +
  labs(title = "Top 10 movies according to IMDB scores",
       y = "IMDB score",
       x = "Movie title") +
  theme(plot.title = element_text(face = "bold", size = 16, hjust = 0),
        plot.subtitle = element_text(face = "italic", size = 11, hjust = 0))
```