---
title: What is the current value of my car? (web scrape basics)
author: Petr Schönbauer
date: '2018-12-10'
slug: web-scraping-basics-what-is-the-current-value-of-my-car
categories: []
tags:
  - regex
  - rvest
  - httr
  - scrape
  - broom
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(rvest)
library(httr)
library(scales)
library(broom)
```

I was wondering what is the current price of my car on the market. One way how to get some at least bit objective value is based on current situation on the market. I´m about to use simple (yet effective) web scraping technique to retrieve data from local best known internet car marketplace.

# First step - how many pages?

As a first step we need to find out on how many pages are the offers related to the same type of my car. By piping the root path to `read_html` and selecting node "a.hide-S"" we can find it out.
```{r}
path <- "https://www.tipcars.com/skoda-rapid/benzin/"
# get max. page
max <- path %>%
  html_session() %>% 
  read_html %>% 
  html_nodes("a.hide-S") %>% 
  html_text()
max
```
..pages, each with 100 cars. Now you may ask how did I know which node to use? And an answer would be simple - using `SelectorGadget` tool. It is pretty cool tool allowing you to select any item on web page of your interest and provide "node" to it. You can find out how to use it [here](https://selectorgadget.com/).

# Load all URLs
Now we can use information about number of pages to create vector of sites we will scrape.
```{r}
url <- paste0(path,"?str=",1:max, "-100")
url
```
To get HTML structure of each page in list is `map` useful function.
```{r}
html_list <- url %>% 
  map(html_session) %>% 
  map(read_html)
```

This simple function extracts text from defined node and html (both input parameters).
```{r}
getText <- function(html, node) {
  html %>%
    html_nodes(node) %>% 
    html_text(trim = TRUE) %>% 
    as.tibble()
}
```

# Map map map 
```{r}
# initiating df with extraction of car prize
df <- html_list %>% map(getText, ".fs-tluste") %>%  map(~.[2:nrow(.),]) %>% bind_rows()
df <- df %>%
  rename(cost = value) %>% 
  # extract engine
  mutate(model = html_list %>% map(getText, ".motor") %>% unlist) %>% 
  # extract mileage
  mutate(mileage = html_list %>% map(getText, ".najeto") %>% unlist) %>% 
  # extract year of production
  mutate(year = html_list %>% map(getText, ".rok_vyroby") %>% unlist)
df
```
Here we have cars of interest (all Škoda Rapid gasoline variants). Let´s decompose it more using a bit of regular expression.

```{r}
final <- df %>% 
  # select numbers only
  mutate(cost = str_extract(cost, "^[0-9]+ [0-9]+")) %>%
  # replace white space
  mutate(cost = str_replace_all(cost, "\\p{WHITE_SPACE}", "")) %>%
  # conversion to double
  mutate(cost = cost %>% as.numeric()) %>%
  # extract power
  mutate(kW = str_extract(model, "[0-9]{2} kW")) %>% 
  # get rid off "kW"
  mutate(kW = str_extract(kW, "[0-9]{2}")) %>% 
  # conversion to factor
  mutate(kW = kW %>% as.factor()) %>% 
  # extract mileage 
  mutate(mileage = str_extract(mileage, "^[0-9]+")) %>%
  # conversion to double
  mutate(mileage = mileage %>% as.numeric() * 1000) %>% 
  select(model, mileage, cost, kW, year)
final
```
Now the data are ready to be plotted.

# Plot plot plot
```{r, warning=FALSE}
final %>% 
  ggplot(aes(x = mileage, y = cost, color = kW)) +
  geom_point() + 
  scale_x_continuous(labels = comma) + 
  scale_y_continuous(labels = comma) + 
  coord_cartesian(xlim = c(0,200000), ylim = c(100000, 400000)) + 
  geom_smooth(method = "lm", se = FALSE) + 
  labs(title = "Škoda Rapid (gasoline)", subtitle = "based on www.tipcars.com", x = "mileage [km]", y = "cost [Kč]")
```

Focused on my car model only:
```{r}
final %>% 
  filter(kW %in% c("77")) %>% 
  
  ggplot(aes(x = mileage, y = cost, color = kW)) +
  geom_point() + 
  scale_x_continuous(labels = comma) + 
  scale_y_continuous(labels = comma) + 
  # coord_cartesian(xlim = c(0,200000), ylim = c(100000, 400000)) + 
  geom_smooth(method = "lm", se = FALSE) + 
  labs(title = "Škoda Rapid 1.2TSI 77kW", subtitle = "based on www.tipcars.com", x = "mileage [km]", y = "cost [Kč]")
```

# So, what is the value? We need a bit of linear regression..
In order to get single value we have to calculate regression coefficients.
```{r}
coeff <- final %>% 
  filter(kW %in% c("77")) %>% 
  lm(cost~mileage, data = .) %>% 
  tidy()
coeff
```
Here we can infer that with each 1km on the road the car looses its value by ~0,5 Kč (~2 Eurocent).

So assuming my car has around 100Tkm mileage its current value is around:

```{r}
my_mileage = 100000
y = coeff$estimate[1] + (my_mileage * coeff$estimate[2])
paste(round(y), "Kč")
```
Or visually:
```{r}
final %>% 
  filter(kW %in% c("77")) %>% 
  
  ggplot(aes(x = mileage, y = cost, color = kW)) +
  geom_point() + 
  scale_x_continuous(labels = comma) + 
  scale_y_continuous(labels = comma) + 
  geom_hline(aes(yintercept = y)) +
  geom_vline(aes(xintercept = my_mileage)) + 
  geom_label(aes(x = my_mileage, y = y,label = paste(round(y), "Kč")), hjust = 0, vjust = 0) + 
  # coord_cartesian(xlim = c(0,200000), ylim = c(100000, 400000)) + 
  geom_smooth(method = "lm", se = FALSE) + 
  labs(title = "Škoda Rapid 1.2TSI 77kW", subtitle = "based on www.tipcars.com", x = "mileage [km]", y = "cost [Kč]")
```

Ok, that´s it. Hope you enjoyed:)