---
title: Temperature change in Czech Republic - part II
author: Petr Schönbauer
date: '2018-12-18'
slug: temperature-change-in-czech-republic-part-ii
tags:
  - broom
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE)
```

```{r, message=FALSE}
library(tidyverse)
library(broom)
library(lubridate)
library(scales)
```

```{r}
df_prec_raw <- readRDS("../../static/data/chmu_srazky.rds")
df_temp_raw <- readRDS("../../static/data/chmu_teploty.rds")
```

```{r}
df_prec_raw
df_temp_raw
```

Merge both dataset together. Check if relevant columns are identical first.
```{r}
map2(df_prec_raw, df_temp_raw, identical)
```

```{r}
df <- bind_cols(df_prec_raw, df_temp_raw) %>% 
  select(year_month, year, month, region, temperature, precipitation) %>% 
  gather(key, value, -year_month, -year, -month, -region) %>% 
  mutate(season = case_when(
    str_detect(month, "Dec|Jan|Feb") ~ "Winter",
    str_detect(month, "Mar|Apr|May") ~ "Spring",
    str_detect(month, "Jun|Jul|Aug") ~ "Autumn",
    str_detect(month, "Sep|Oct|Nov") ~ "Summer"))
df
```

# What was the question?

## Claim - "The precipitation in Czech Republic decreases over last decades."

```{r}
df %>% 
  
  filter(key == "precipitation") %>% 
  filter(region == "Česká republika") %>% 
  
  ggplot(aes(year_month, value)) + 
  geom_col() + 
  geom_smooth(method = "lm") +
  scale_x_datetime(labels = date_format("%Y"), date_breaks = "1 year") + 
  theme(axis.text.x = element_text(angle = 75, hjust = 1)) + 
  labs(y = "precipitation [mm]", x = "year") +
  facet_wrap(~region, scales = "free")
```

Nope (on average).

## Claim - "It rains less and less every summer in Czech Republic."

```{r, message=FALSE}
df %>% 
  
  filter(key == "precipitation") %>% 
  filter(region == "Česká republika") %>% 
  
  ggplot(aes(year_month, value)) + 
  geom_col() + 
  geom_smooth(method = "lm") +
  # coord_flip() + 
  scale_x_datetime(labels = date_format("%Y"), date_breaks = "1 year") + 
  theme(axis.text.x = element_text(angle = 75, hjust = 1)) + 
  labs(y = "precipitation [mm]", x = "year") +
  facet_wrap(~season, scales = "free")
```

Well, seems to be quite stable, no drama here.

## Claim - "The temperature in Czech Republic increases over last decades."

```{r, message=FALSE}
df %>% 
  
  filter(key == "temperature") %>% 
  filter(region == "Česká republika") %>% 
  
  ggplot(aes(year_month, value)) + 
  geom_col() + 
  geom_smooth(method = "lm") +
  scale_x_datetime(labels = date_format("%Y"), date_breaks = "1 year") + 
  theme(axis.text.x = element_text(angle = 75, hjust = 1)) + 
  labs(y = "°C", x = "year") + 
  facet_wrap(~region)
```

Yes, it does increase. On average by:

```{r}
df %>% 
  
  filter(key == "temperature") %>% 
  filter(region == "Česká republika") %>% 
  
  mutate(year = year %>% as.integer()) %>% 

  group_by(year) %>% 
  summarise(temp_avg_year = mean(value)) %>% 
  
  lm(temp_avg_year~year, data = .) %>%
  tidy() %>%
  pull(estimate) %>% 
  .[[2]] %>% 
  round(3)
```

..°C per year.